Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages

### 2018-01-15 paper/manuscript.tex
Start adding to Methods

\subsection*{Creation of MDCK-SIAT1-TMPRSS2 cell line}
The human TMPRSS2 cDNA ORF was ordered from OriGene (NM\_005656), PCR amplified, and cloned into a pHAGE2 lentiviral vector under an EF1$\alpha$-Int promoter and attached to mCherry through an IRES...etc etc
\comment{Need to look at Katie's notebooks for this...}

\subsection*{Generation of HA codon mutant plasmid libraries}
Recombinant A/Perth/16/2009 (HA, NA) $\times$ A/Puerto Rico/8/1934 influenza virus, NIB-64, NR-41803 was ordered from BEI Resources, NIAID, NIH. 
Bulk RNA from the viral sample was extracted using the QIAamp Viral RNA Mini Kit (QIAGEN) according to manufacturer's instructions.
The Perth/2009 HA and NA genes were then reverse transcribed, PCR amplified, and cloned into the pHW2000~\citep{hoffmann2000dna} and pICR2 \comment{cite?} plasmid backbones.

The codon-mutant libraries were generated using a PCR-based approach described in~\cite{bloom2014experimentally,dingens2017comprehensive}.

\subsection*{Generation and passaging of mutant viruses}
The mutant virus libraries were generated and passaged using the approach described in~\cite{doud2016accurate} with several modifications.

\subsection*{Barcoded subamplicon sequencing}

\subsection*{Analysis of deep sequencing data}

\subsection*{Inference of phylogenetic trees}
\comment{We downloaded X sequences from the Influenza Virus Resource ?.... etc. inferred the tree, ancestral state reconstruction, visualized the tree...}
To parse out trunk mutations from side branch mutations, we first defined a set of recent nodes sampled on or after Jan. 1, 2017, and traced these nodes back to their most recent common ancestor.

>>>>

\subsection*{Creation of MDCK-SIAT1-TMPRSS2 cell line}
The human TMPRSS2 cDNA ORF was ordered from OriGene (NM\_005656), PCR amplified, and cloned into a pHAGE2 lentiviral vector under an EF1$\alpha$-Int promoter and attached to mCherry through an IRES. 
We generated lentiviruses..etc etc transduced MDCK-SIAT1 cells, sorted by fluorescence activated cell sorting
\comment{Need to look at Katie's notebooks for this...}

\subsection*{Generation of HA codon mutant plasmid libraries}
Recombinant A/Perth/16/2009 (HA, NA) $\times$ A/Puerto Rico/8/1934 influenza virus, NIB-64, NR-41803 was ordered from BEI Resources, NIAID, NIH. 
Bulk RNA from the viral sample was extracted using the QIAamp Viral RNA Mini Kit (QIAGEN) according to manufacturer's instructions.
The Perth/2009 HA and NA genes were then reverse transcribed using the SuperScript III First-Strand Synthesis SuperMix Kit (Invitrogen), PCR amplified, and cloned into the pHW2000~\citep{hoffmann2000dna} and pICR2 \comment{cite?} plasmid backbones by In-Fusion.

To generate wildtype Perth/2009 H3N2 virus, we transfected a co-culture of 293T and MDCK-SIAT1-TMPRSS2 cells with reverse genetics plasmids encoding the Perth/2009 HA and NA, WSN/1933 internal genes, and the TMPRSS2 lentiviral vector.
\comment{add more details about reverse genetics?}
The viral supernatant was harvested at 72 hours post-transfection, clarified at 1200 rpm for 5 min, and 0.01 to 1.0 ul of the supernatant was passaged onto $4 \times 10^{5}$ MDCK-SIAT1-TMPRSS2 cells in six-well plates for a total of six serial passages.
After the final passage, RNA was extracted from 140 ul of the viral supernatant using the QIAamp Viral RNA Mini Kit.
The HA and NA genes were reverse transcribed, PCR amplified, and sequenced by Sanger sequencing.
Two mutations, G78D and T212I, arose after serial passaging. 
The Perth/2009 HA carrying these two mutations were cloned into the pHW2000~\citep{hoffmann2000dna} and pICR2 \comment{cite?} plasmid backbones.

The codon-mutant libraries were generated in the Perth/2009 HA-G78D-T212I background using a PCR-based approach described in~\cite{bloom2014experimentally,dingens2017comprehensive} using two rounds of mutagenesis.
The codon mutations were introduced into three independent replicates of the Perth/2009 HA variant.
The mutant variants were then cloned into the pICR2 vector by digestion with BsmBI restriction sites, ligation by T4 DNA ligase, and electroporation and transformation into ElectroMAX DH10B competent cells (Invitrogen; 18290015). 
We pooled over 6 million transformants for each replicate, cultured in liquid LB + ampicillin at 37$^{\circ}$C for 3 h with shaking, and maxiprepped.

\subsection*{Generation and passaging of mutant viruses}
The mutant virus libraries were generated and passaged using the approach described in~\cite{doud2016accurate} with several modifications.
We used the HA-deficient WSN/1933 helper virus generated in~\cite{doud2016accurate}.
Briefly, these helper viruses were generated by transfecting seven WSN/1933 non-HA reverse-genetics plasmids plus a WSN/1933 HA protein expression plasmid into a coculture of 293T and MDCK-SIAT1-EF1$\alpha$-WSN-HA cells engineered to constitutively express the WSN/1933 HA.
The transfection supernatant was then expanded by passaging onto a monolayer of MDCK-SIAT1-EF1$\alpha$-WSN-HA cells, and the viral supernatant was then harvested and aliquots were frozen at -80$^{\circ}$C.

To generate the mutant virus libraries, we transfected $5 \times 10^5$ MDCK-SIAT1-TMPRSS2 cells in suspension with 937.5 ng each of four protein expression plasmids encoding the ribonucleoprotein complex (HDM-Nan95-PA, HDM-Nan95-PB1, HDM-Nan95-PB2, and HDM-Aichi68-NP)~\cite{gong2013stability}, and 1250 ng of one of the three pICR2-mutant-HA libraries (or the wildtype pICR2-Perth2009-HA-G78D-T212I control) using Lipofectamine3000 (ThermoFisher;L3000008).
We allowed the transfected cells to adhere in 6-well plates, then four hours later changed the media to D10 media (DMEM supplemented with 10\% heat-inactivated FBS, 2 mM L-glutamine, 100 U of penicillin/mL, and 100 $\mu$g of streptomycin/mL).
Eighteen hours after transfection, we infected the cells with WSN/1933 HA-deficient helper virus by preparing an inoculum of 500 TCID$_{50}$ per $\mu$L in influenza growth media (Opti-MEM supplemented with 0.01\% heat-inactivated FBS, 0.3\% BSA, 100 U of penicillin/mL, 100 $\mu$g of streptomycin/mL, and 100 $\mu$g of calcium chloride/mL), and aspirating the D10 media from the cells, and adding 2 mL of the helper-virus inoculum to each well.
After three hours, we changed the media to fresh influenza growth media.
Twenty-four hours after helper-virus infection, we harvested the viral supernatants for each replicate, froze aliquots at -80$^{\circ}$C, and titered in MDCK-SIAT1-TMPRSS2 cells.

We passaged over $9 \times 10^5$ TCID$_{50}$ of the transfection supernatants at an MOI of 0.0035 TCID$_{50}$ per cell.
To passage, we plated $4.6 \times 10^6$ MDCK-SIAT1-TMPRSS2 cells per dish in 15-cm dishes in D10 media, and allowed the cells to grow for 24 hours, at which the cells reached a density of $\sim 1.7 \times 10^7$ cells per dish.
We then replaced the media of each dish with 25 mL of an inoculum of 2.5 TCID$_{50}$ of virus per $\mu$L.
We collected viral supernatant for sequencing 48 hours post-infection.

\subsection*{Barcoded subamplicon sequencing}
To extract viral RNA from the three replicate HA virus libraries and the wildtype HA virus, we first ultracentrifuged 24 mL of the supernatant at 22,000 rpm for 1.5 h at 4$^{\circ}$C in a Beckman Coulter SW28 rotor.
We then extracted the RNA using the Qiagen RNeasy Mini Kit by resuspending the viral pellet in 400 $\mu$L of buffer RLT supplemented with $\beta$-mercaptoethanol, pipetting 30 times, transferring the liquid to a microcentrifuge tube, adding 600 $\mu$L 70\% ethanol, and proceeding with the rest of the extraction according to manufacturer's instructions.
The HA gene was then reverse-transcribed with AccuScript Reverse Transcriptase (Agilent 200820) using the primers P09-HA-For (5'-AGCAAAAGCAGGGGATAATTCTATTAATC-3') and P09-HA-Rev (5'-AGTAGAAACAAGGGTGTTTTTAATTACTAATACAC-3').

We generated the HA PCR amplicons for each of the three plasmid libraries, the three virus libraries, one wildtype plasmid, and one wildtype virus samples using KOD Hot Start Master Mix (EMD Millipore;71842) according to the PCR reaction mixture and cycling conditions described in~\cite{bloom2014experimentally} and the P09-HA-For and P09-HA-Rev primers.
\comment{used the same conditions as Mike - cite his Viruses paper to be succinct?}.
We performed sequencing on a lane of a flow cell of an Illumina HiSeq 2500 using 2 $\times$ 250 bp paired-end reads in rapid-run mode.

\subsection*{Analysis of deep sequencing data}
We used dms\_tools2...

\subsection*{Inference of phylogenetic trees}
\comment{John to add methods here. We downloaded X sequences from the Influenza Virus Resource ?.... etc. inferred the tree, ancestral state reconstruction, visualized the tree...}
To parse out trunk mutations from side branch mutations, we first defined a set of recent nodes sampled on or after Jan. 1, 2017, and traced these nodes back to their most recent common ancestor.
